Displacement of fishing effort by Large Scale Marine Protected Areas

### 2018-10-17 docs/Manuscript.tex
Updated manuscript intro and methods

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Displacement of fishing effort by Large Scale Marine Protected Areas}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-11}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers

@@

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from extractive activities such as fishing. Current international
goals aim to ``effectively protect 10\% of the ocean environments by
2020''. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018b}. A significant
part of this rapid increase can be attributed to the designation of a
small number of Large Scale Marine Protected Areas {[}LSMPAs;
\citep{gray_2017}{]}. These are defined as MPAS with at least 250,000
km\textsuperscript{2} in extension \citep{toonen_2013}, and are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017}.

Due to weak property rights, limited habitat transformation, and
potentially lower management costs, pelagic MPAs provide a great
opportunity to safeguard the oceans \citet{game_2009}{[}\^{}d{]}. A
growing body of literature has also shown that closing the high seas to
all fishing -effectively, turning them into a LSMPA- could increase
fishery yields and profitability of fisheries, without negligible costs
to food security
\citep{white_2014,sumaila_2015,sala_2018,schiller_2018}. However, as
with customary MPAs, it is important that we understand the
socioeconomic implications of management interventions.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. LSMPAs were erroneously assumed to have little social
implications due to their remoteness \citep{agardy_2011,gray_2017}.
However, the anticipation of a LSMPA can lead to preemptive overfishing,
which can erode or delay the expected benefits of the intervention
\citep{mcdermott_2018}. LSMPAs have received great attention in terms of
governance and enforcement, but are yet to be the focus of economic
analyses \citep{gray_2017}. For example, \citet{mcdermott_2018} show
that fishing effort within the Phoenix Islands Marine Protected Area
(PIPA) is effectively reduced after implementation, and describe changes
in fishing behavior pre-implementation. \citet{cabral_2017} analyse the
redistribution of fishing and non-fishing vessels following the
implementation of MPAs in California, and find that responses are
idiosyncratic; commercial dive boats follow a fishing-the-line pattern,
while some fishing boats follow an ideal free distribution. The way in
which fishers react to a spatial closure can have major implications in
its outcome \citep{hilborn_2006,krueck_2017,viana_2017}. This highlights
the need to understand how fishers react to the implementation of a
LSMPA, and fishing effort changes and is spatially redistributed.

The main objective of this paper is to identify how fishers adapt to the
implementation of LSMPAs. We combine novel vessel tracking technologies
and causal inference techniques to identify behavioral changes of
fishing vessels due to the implementation of PIPA. We focus on fishing
hours and distance traveled as outcome variables that fishers might
adjust following implementation of a LSMPA in an impact-evaluation
fashion. Additionally, we evaluate the spatial redistribution of fishing
effort that existed within PIPA before its implementation. This work
provides novel empirical insights into fisher's responses to the
implementation of LSMPAs, and can help guide future interventions.

@@

Our data contain over 45 million individual AIS messages (\emph{i.e.}
positions) for 371 purse seiners and longliners. A total of 233 vessels
have fished within PIPA waters; 217 did so at least once before 2015.
However, not all vessels continued to fish elsewhere after PIPA
implementation: 34 vessels have no recorded AIS messages after
2015\footnote{The 34 missing vessels might have exited the fishery, been
  decommissioned or sold (therefore changing their AIS and mmsi), or
  turned off their AIS transmitters. In either case, we are not able to
  observe these.}, leaving us with 176 vessels that fished inside PIPA
before its implementation, and continued to fish elsewhere afterward.
New vessels might have also entered the fishery after PIPA closure, and
were likely not exposed to the policy intervention in the pre-treatment
period. To account for this, we identify a subset of vessels which we
track since before the implementation of PIPA, and categorize them as
treated or control vessels. Our treatment and control groups are defined
as follows.

The treatment group contains all vessels (n = 176) that fished within
PIPA at least once before the closure, and that continued to fish
elsewhere afterwards. Vessels in the control group meet all three of the
following conditions: i) vessels never fished within PIPA waters, ii)
vessels belong to other PNA countries, and iii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure. For each vessel meeting these characteristics, we calculate
their total monthly fishing hours. Figure \ref{fig:baci_strict} provides
a visual representation of the vessel-level fishing events that make up
each group through time. Table \ref{tab:baci_n_s} shows the number of
vessels following a BACI design, as well as the fishing hours, before
and after PIPA.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-20-1.pdf}
\caption{\label{fig:unnamed-chunk-20}\label{fig:baci_strict}Stream of
fishing events by vessels through time. Each line represents a vessel,
with dots indicating months with fishing activity and colors indicating
the pre and post periods.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-21}\label{tab:baci_n_s}Number of fishing vessels and fishing hours by gear and treatment group before and after PIPA.}
\centering
\begin{tabular}[t]{llrrrr}
\toprule
Gear & Treatment & n & Before & After & Change (A / B)\\
\midrule
drifting\_longlines & FALSE & 77 & 39.03181 & 37.79044 & 0.9681958\\
drifting\_longlines & TRUE & 115 & 41.62659 & 40.43857 & 0.9714600\\
purse\_seines & FALSE & 26 & 11.73804 & 20.13247 & 1.7151474\\
purse\_seines & TRUE & 61 & 10.66939 & 18.36289 & 1.7210820\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. Our variables of interests are fishing effort,
indicated by total fishing hours per month, and distance traveled (Km)
on every fishing trip. We compare fishing hours\footnote{And soon,
  distance} before and after the implementation of PIPA using a
Difference-in-Differences approach, where we track the variable of
interest for vessels that used to fish inside PIPA and vessels that
never fished inside PIPA, before and after PIPA implementation. Our
specification is the following:

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \mu_1Y_t + \mu_2Y_t^2 + \phi_t + \gamma_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\mu_1\) and \(\mu_2\) are coefficients
for a second order polynomial for years (\(Y_t\)), while \(\phi_t\) and
\(\gamma_i\) represent month-, and flag-level dummies that account for
seasonality or country-level management interventions\footnote{An
  earlier specification included years as a dummy variable. Such results
  are included in the appendix, but are similar to the ones found under
  current specification.}.

Our second part of the analyses focuses on the redistribution of fishing
effort. In other words, identifying where do vessels that used to fish
inside PIPA go after its establishment. We calculate the monthly
relative distribution of fishing hours by all treated vessels across all
fished EEZs and the high seas. These trends are shown in Figure
\ref{fig:redist_trend_ps}, and the relative temporal change is presented
in Table \ref{tab:ba_disp}. EEZs that had sporadic fishing events were
pooled into a group of ``others'', leaving us with a total of n = 23 and
n = 0\footnote{This number is likely to change upon finalizing the
  spatial analysis of longliners, which is currently running.} spatially
defined regions (\emph{i.e.} EEZs, High Seas, ``other EEZs'', and PIPA)
for purse seiners and longliners, respectively.

To evaluate this change in effort allocation, we regress our variable of
interest (\emph{i.e.}fishing hours) on the interaction between a dummy
variable indicating the policy intervention and a dummy variable for
countries. This gives us the by-country change in proportional
allocation of fishing effort:

\[
y_{i,t} = \alpha + \beta_1Post_t + \beta_{2,i}Country_i + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries, interpreted as individual EEZs, the high seas, and a
group of ``other EEZs''. Our parameter of interest is \(\beta_{3,i}\),
which captures the country-level change in proportional fishing effort.

@@

@@

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-25-1.pdf}
\caption{\label{fig:unnamed-chunk-25}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

@@

Longliners show a similar pattern of effort reduction. However, the
magnitude of the \(\beta_3\) coefficient is smaller (ranging from
\(\beta_3 = -9.851\) to \(\beta_3 = -14.850\)) and not significant
across all model specifications. This, along with higher standard error
values suggest that longliners have a smaller and more variable response
to the implementation of LSMPAs.

@@

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-29-1.pdf}
\caption{\label{fig:unnamed-chunk-29}\label{fig:redist_trend_ps}Monthly
relative allocation of fishing effort by PIPA-fishing vessels before and
after PIPA for 9 EEZs, PIPA, the high seas and `other EEZs'.}
\end{figure}

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Displacement of fishing effort by Large Scale Marine Protected Areas}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-16}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

@@

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}. A significant part
of this rapid increase can be attributed to the designation of a small
number of Large Scale Marine Protected Areas (LSMPAs;
\citet{boonzaier_2016,singleton_2014,gray_2017}). These are defined as
MPAS larger than 250,000 Km\textsuperscript{2} in extension (Fig.
\ref{fig:LSMPAs_map}), and are often implemented in the pelagic
environment, where the dominant human activity is industrial fishing
\citep{toonen_2013,gray_2017,kroodsma_2018}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. Moreover, the world has seen a widespread increase in
the implementation of LSMPAs, likely as a way to rapidly achieve the
10\% target \citep{boonzaier_2016,alger_2017,sala_2018}. However, as
with customary MPAs, it is important that we understand the
socioeconomic implications of management interventions that result in
the spatial displacement of fishing effort. They \textbf{key gap in the
literature is X}. We adress it here by using new stuff and find results
that are of relevance to policy makers and advance our knowledge, which
should be incorporated when designing and planning for LSMPAs.

Due to weak property rights, limited habitat transformation, and
potentially lower management costs, pelagic MPAs provide an opportunity
to safeguard the oceans \citet{game_2009}. A growing body of literature
has shown that closing the high seas to all fishing could increase
fishery yields and profitability of fisheries, with negligible costs to
food security \citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}.

The early literature on LSMPAs focused on the inherent difficulties that
come with a pelagic (\emph{i.e.} open water) environment, where
organisms typically have large home ranges. \citet{game_2009} suggest
that most of the challenges can be overcome with the incorporation of
technology, in what then became known as Dynamic Ocean Management
\citep{maxwell_2015}. \citet{kaplan_2009} claim that very large MPAs
would result in excessive opportunity costs and that these would be
difficult to enforce. \citet{toonen_2013} show that just a small number
of LSMPAs made up 80\% of the managed areas in the ocean in 2013, and
that these can help achieve the CBD Aichi targets faster.
\citet{singleton_2014} provide an objective discussion of pros and cons
of LSMPAs.

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these human
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protectes two
thrids of the tagged grey reef sharks by effectively excluding fishing
effort. No studies have evaluated the displacement of fishing effort due
to LSMPA implementation.

The exclusion of these LSMPAs is likely to change fisher's behavior.
Theoretical models of fishing effort redistribution range from the
simplistic assumption that effort inside the bounded region disappears,
to spatially explicit models that reallocate fishing effort based on
habitat characteristics, presence of other vessels, and expected returns
\citep{smith_2003,hilborn_2006}. However, these focus on the long term
optimal equilibrium and redistribution of fishing effort may not be
optimal, especially over the first years \citep{stevenson_2013}. The few
empirical works suggest that resource users may show idiosyncratic
responses. For example, \citet{stevenson_2013} show that a network of
MPAs displaced fishing effort farther away from ports resulting in
higher \emph{perceived} costs, but that catch per unit effort also
increased. \citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of MPAs in California,
and find that commercial dive boats follow a fishing-the-line pattern,
while some fishing boats follow an ideal free distribution. The way in
which fishers react to a spatial closure can have major implications in
its outcome \citep{smith_2003,hilborn_2006} highlighting the need to
understand how fishers react to the implementation of LSMPAs, and
fishing effort changes and is spatially redistributed.

The main objective of this paper is to identify behavioral responses to
the implementation of LSMPAs. We combine novel vessel tracking
technologies and causal inference techniques to identify behavioral
changes of fishing vessels due to the implementation of PIPA. We focus
on fishing hours and distance traveled as outcome variables that fishers
might adjust following implementation of a LSMPA in an impact-evaluation
fashion. Additionally, we evaluate the spatial redistribution of fishing
effort that existed within PIPA before its implementation. This work
provides novel empirical insights into fisher's responses to the
implementation of LSMPAs, and can help guide future interventions. Our
work is novel in the sense that it provides empirical evidence of the
effect of large-acale Marine Protected Areas in fishing behavior and
distribution.

The next sections are as follows: Section \ref{background} provides an
overview of the Nauru Agreement and associated countries, a description
of the fleet that operates in the region, and a brief history of PIPA.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other LSMPAs and discusses our
results.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

\hypertarget{background}{%
\section{Background}\label{background}}

The Nauru Agreement was established in 1982 by Pacific island nations to
manage their important tuna resources. PNA Members include Federated
States of Micronesia, Kiribati, Marshall Islands, Nauru, Palau, Papua
New Guinea, Solomon Islands, and Tuvalu.The Nauru Agreement regulated
access of foreign vessels (\emph{i.e.} those from non-PNA countries).
Holding \textasciitilde{}80\% of the historical purse seining grounds
within their Exclusive Economic Zones, PNA countries gained bargaining
power when providing access to foreign fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisors recomend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. A solid line indicates
countries that belong to the PNA, while a dashed line indicates all
others. A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. The pacific region
has historically accounted for a large portion of tuna species
production \citep{aqorau_1997}. Today, the PNA controls close to 50\% of
the global skipjack tuna production \citep{pna_website_2018}. A large
portion of these catches derive from purse seine vessels licensed under
the VDS. Fishing vessels from Australia, New Zeland, China, France,
Korea, Japan, the Philippines, Taiwan, and the United States participate
in the purse-seining VDS.

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 Km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. The implementation of such a large
conservation area in one of the most important fishing regions in the
world provides a great opportunity to evaluate the behavioral responses
and redistribution of fishing effort by vessels that used to fish there.
\citet{mcdermott_2018} showed that fishing effort within the Phoenix
Islands Marine Protected Area (PIPA) increased between the anouncement
(September 1\textsuperscript{st}, 2014) and its implementation in
Janyary 1\textsuperscript{st}, 2015. Likewise, they demonstrate that
fishing effort is effectively reduced after implementation. To this, we
pose two questions: How do individual vessels respond to the sudden
exclusion of such a big area? And where did all the vessels go? In the
next sections we describe the data and methods used to answer this
questions.

@@

Our data contain over 45 million individual AIS messages (\emph{i.e.}
positions) for 371 fishing vessels that at some point fished in PNA
waters. A total of 233 vessels have fished within PIPA waters; 217 did
so at least once before 2015. However, not all vessels continued to fish
elsewhere after PIPA implementation: 34 vessels have no recorded AIS
messages after 2015, leaving us with 176 vessels that fished inside PIPA
before its implementation, and continued to fish elsewhere
afterward\footnote{The 34 missing vessels might have exited the fishery, been decommissioned or sold (therefore changing their AIS and mmsi), or turned off their AIS transmitters. In either case, we are not able to observe these.}.
From these, we focus on purse seiners -the most important for PNA
countries- and are left with 61
vessels\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
New vessels might have also entered the fishery after PIPA closure, and
were likely not exposed to the policy intervention in the pre-treatment
period. To account for this, we identify a subset of vessels which we
track since before the implementation of PIPA, and categorize them as
treated or control vessels. Our treatment and control groups are defined
as follows.

The treatment group contains all purse seiners (n = 61) that fished
within PIPA at least once before the anouncement (09/01/2014), and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure. This definition of control vessels provides us with 26 purse
seiners that never fished inside PIPA but fished in other PNA waters
before and after PIPA implementation. During our causal identification,
we also use two more strict definitions of control groups: one with only
vessels that belong to PNA countries (n = 7), and one that excludes
Chinese vessels (n = 21). Our definition of treatment and control groups
leaves us with 61 and 21 treated and control vessels, which have just
over 22 million observations where about 22\% are identified as fishing
events. For each of the remainig vessels, we calculate their total daily
fishing hours and obtain a pannel data with 38,200 observations. Table
\ref{tab:baci_n_s} shows the number of vessels following a BACI design,
as well as the fishing hours, before and after PIPA. Figure
\ref{fig:baci_strict} provides a visual representation of the
vessel-level fishing events that make up each group through time.

@@

\caption{\label{tab:unnamed-chunk-6}\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Group & n & Before & After & Change (A / B)\\
\midrule
Control & 26 & 11.74 & 20.13 & 1.72\\
Treatment & 61 & 10.67 & 18.36 & 1.72\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. Our variables of interests are daily fishing
hours, and daily distance traveled (Km). We compare our variable of
interest before and after the implementation of PIPA using a
Difference-in-Differences approach, where we track the variable of
interest for vessels that used to fish inside PIPA and vessels that
never fished inside PIPA, before and after PIPA implementation. Our
specification is the following:

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month-, and flag-level dummies that account for seasonality or
country-level management
interventions\footnote{We test for other more complex specifications that interact a quarterly or year-month variable with the treatment group and find qualitativel the same results.}

Our second part of the analyses focuses on the redistribution of fishing
effort. In other words, identifying where do vessels that used to fish
inside PIPA go after its establishment. We discretize spatial units by
using a polygon for PIPA
\footnote{we would expect to see a decrease here} and distinct spatial
units for each EEZ of each country. Some vessels might shift from EEZs
into the High Seas, but we are interested in knowing \emph{where} in the
High Seas, so we incorporate additional regions by using a 1 degree
buffer of the High Seas arround each of the EEZ regions. The rest of the
high seas are merged into a single spatial unit. For example, if we were
to do this only for Kiribati, we would have 8 spatial units: PIPA, three
EEZs, three 1-degree buffers of High Seas arround each EEZs, and the
rest of the High Seas. Whenever the buffers overlapped between
themselves, we randocmply clipped one on to the other. EEZs that had
sporadic fishing events were pooled into a group of ``others''.

To evaluate this change in effort allocation, we regress our variable of
interest (\emph{i.e.} fishing hours) on the interaction between a dummy
variable indicating the policy intervention and a dummy variable for
countries. This gives us the by-country change in proportional
allocation of fishing effort:

\[
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries for the spatial units defined above. Our parameter of
interest is \(\beta_{1,i}\), which captures the country-level change in
proportional fishing effort.

@@

\caption{\label{tab:unnamed-chunk-9}\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.}
\centering
\begin{tabular}[t]{lr}
\toprule
country & change\\
\midrule
EEZ COK 1 & 0.53\\
EEZ FSM 1 & 0.81\\
EEZ KIR 1 & -0.16\\
EEZ KIR 2 & 4.50\\
EEZ KIR 3 & -2.79\\
\addlinespace
EEZ MHL 1 & -0.44\\
EEZ NRU 1 & 0.07\\
EEZ PNG 2 & -9.07\\
EEZ SLB 1 & 3.18\\
EEZ TUV 1 & 1.27\\
\addlinespace
HS & 3.93\\
HS COK 1 & 0.06\\
HS FSM 1 & 0.01\\
HS KIR 1 & 2.96\\
HS KIR 2 & 1.07\\
\addlinespace
HS KIR 3 & 2.53\\
HS MHL 1 & 0.11\\
HS MUS 1 & 0.00\\
HS NRU 1 & -0.08\\
HS PNG 2 & -0.20\\
\addlinespace
HS SLB 1 & -0.02\\
HS TUV 1 & 0.01\\
PIPA PIPA 1 & -8.25\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{\label{fig:unnamed-chunk-10}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{\label{fig:unnamed-chunk-11}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
\caption{\label{fig:unnamed-chunk-15}\label{fig:redist_trend_ps}Monthly
relative allocation of fishing effort by PIPA-fishing vessels before and
after PIPA for 9 EEZs, PIPA, the high seas and `other EEZs'.}
\end{figure}
